Concurrent Programming with CXL

### 2020-06-19 book/paper.tex
Starting to make machine code easier to read

\begin{figure}
\begin{code}
\begin{verbatim}
Up.cxl:1 def incrementer(self):
    0 Jump 39
    1 Frame incrementer ['self'] 12
Up.cxl:2     count = count + 1;
    2 Push 1
    3 PushAddress count
    4 Load 1
    5 2-ary +
    6 PushAddress count
    7 Store 1
Up.cxl:3     done[self] = True;
    8 Push True
    9 PushVar self
   10 PushAddress done
   11 Store 2
   12 Return
\end{verbatim}
\end{code}
\caption{The first part of the CVM bytecode corresponding to Figure~\ref{fig:inc}.}
\label{fig:inccode}
\end{figure}

\begin{figure}
\begin{code}
\begin{verbatim}
CXL Assertion failed ('assert', 'Up.cxl', 9, 5) 1
#states = 79
==== Safety violation ====
__init__/() [0,39-56]       57 dict{ .count:0, .done:dict{ 0:False, 1:False } }
incrementer/0 [1-6]          7 dict{ .count:0, .done:dict{ 0:False, 1:False } }
incrementer/1 [1-11]        12 dict{ .count:1, .done:dict{ 0:False, 1:True  } }
incrementer/0 [7-11]        12 dict{ .count:1, .done:dict{ 0:True,  1:True  } }
main/() [14-19,22-27,29-36] 37 dict{ .count:1, .done:dict{ 0:True,  1:True  } }
\end{verbatim}
\end{code}
\caption{The output of running Figure~\ref{fig:inc}.}
\label{fig:incoutput}
\end{figure}

@@

\begin{enumerate} \setcounter{enumi}{1}
\item the \texttt{Push} instruction pushes the constant 1
onto the stack of the process.
\item the \texttt{PushAddress} instruction pushes the address of
the shared variable \texttt{count} onto the stack.
\item The \texttt{Load} instruction pops the address of the
\texttt{count} variable and then pushes the value of the
\texttt{count} variable onto the stack.
\item \texttt{2-ary} is a \texttt{+} operation with 2 arguments.
It pops two values from the stack (1 and the value of \texttt{count}),
adds them and pushes the result back onto the stack.
\item The \texttt{PushAddress} instruction pushes the address of the \texttt{count}
variable onto the stack.
\item The \texttt{Store} instruction pops the address of a variable and pops
a CXL value (the sum of the \texttt{count} variable and 1) and updates the
\texttt{count} variable.
\end{enumerate}

@@

If we look at the middle three rows, we see that:
\begin{enumerate}
\item Process 0 executed instructions 1 through 7, loading the value of
\texttt{count} but stopping just before storing 1 into \texttt{count};
\item Process 1 executed instructions 1 through 12, storing 1 into
into \texttt{count} and storing \texttt{True} into \texttt{done[1]};
\item Process 0 continues execution, storing value 1 into \texttt{count}
and storing \texttt{True} into \texttt{done[0]}.
\end{enumerate}

@@

\vspace{1em}
\begin{tabular}{|l|l|}
\hline
Address & compute address \\
\hline
Apply & pop $m$ and $i$ and apply $i$ to $m$, pushing a value \\
\hline
Assert & pop $b$ and check that it is \texttt{True} \\
\hline
AtomicInc & increment the atomic counter of this context \\
\hline
AtomicDec & decrement the atomic counter of this context \\
\hline
Continue & no-op (but causes a context switch) \\
\hline
Choose & choose an element from the set on top of the stack \\
\hline
Del & delete a shared variable \\
\hline
DelVar $v$ & delete process variable $v$ \\
\hline
Dict & pop $n$ and $n$ key/value pairs and push a dictionary \\
\hline
Dup & duplicate the top element of the stack \\
\hline
Frame $m$ $a$ $r$& start method $m$ with arguments $a$,
initializing variables.  Last instruction is at $r$ \\
\hline
Go & pop context and value, push value on context's stack, and add to context bag \\
\hline
Jump $p$ & set program counter to $p$ \\
\hline
JumpCond $e$ $p$ & pop expression and, if equal to $e$, set program counter to $p$ \\
\hline
Load & load a shared variable and push it onto the stack \\
\hline
LoadVar $v$ & load process variable $v$ and push it onto the stack \\
\hline
Nary \textit{op} $n$ & apply $n$-ary operator \textit{op} to the top $n$ elements on the stack \\
\hline
Pop & pop a value of the stack and discard it \\
\hline
Push $c$ & push constant $c$ onto the stack \\
\hline
PushAddress $v$ & push the address of shared variable $v$ \\
\hline
Return & pop return address, push \texttt{result}, and restore program counter \\
\hline
Set & pop $n$ and $n$ elements and push a set of the elements \\
\hline
Spawn & pop tag, argument, and method and spawn a new context \\
\hline
Split & pop set and push minimum and remainder \\
\hline
Stop & save context into shared variable and remove from context bag \\
\hline
Store & pop a value from the stack and store it in a shared variable \\
\hline
StoreVar $v$ & pop a value from the stack and store it in process variable $v$ \\
\hline
\end{tabular}
\vspace{1em}

>>>>

\begin{figure}
\begin{code}
\begin{verbatim}
Up.cxl:1 def incrementer(self):
   0 Jump 32
   1 Frame incrementer ['self'] 9
Up.cxl:2     count = count + 1;
   2 Push 1
   3 Load count
   4 2-ary +
   5 Store count
Up.cxl:3     done[self] = True;
   6 Push True
   7 PushVar self
   8 Store done 1
   9 Return
\end{verbatim}
\end{code}
\caption{The first part of the CVM bytecode corresponding to Figure~\ref{fig:inc}.}
\label{fig:inccode}
\end{figure}

\begin{figure}
\begin{code}
\begin{verbatim}
   ctx = __init__/() pc =   0 #states = 1 diameter = 0 queue = 0
CXL Assertion failed ('assert', 'Up.cxl', 9, 5) 1
#states = 79
==== Safety violation ====
__init__/() [0,32-47]       48 dict{ .count:0, .done:dict{ 0:False, 1:False } }
incrementer/0 [1-4]          5 dict{ .count:0, .done:dict{ 0:False, 1:False } }
incrementer/1 [1-8]          9 dict{ .count:1, .done:dict{ 0:False, 1:True  } }
incrementer/0 [5-8]          9 dict{ .count:1, .done:dict{ 0:True,  1:True  } }
main/() [11-15,18-22,24-29] 29 dict{ .count:1, .done:dict{ 0:True,  1:True  } }
\end{verbatim}
\end{code}
\caption{The output of running Figure~\ref{fig:inc}.}
\label{fig:incoutput}
\end{figure}

@@

\begin{enumerate} \setcounter{enumi}{1}
\item the \texttt{Push} instruction pushes the constant 1
onto the stack of the process.
\item The \texttt{Load} instruction pushes the value of the
\texttt{count} variable onto the stack.
\item \texttt{2-ary} is a \texttt{+} operation with 2 arguments.
It pops two values from the stack (1 and the value of \texttt{count}),
adds them and pushes the result back onto the stack.
\item The \texttt{Store} instruction pops pops
a CXL value (the sum of the \texttt{count} variable and 1) and
stores it in the \texttt{count} variable.
\end{enumerate}

@@

If we look at the middle three rows, we see that:
\begin{itemize}
\item Process 0 executed instructions 1 through 4, loading the value of
\texttt{count} but stopping just before storing 1 into \texttt{count};
\item Process 1 executed instructions 1 through 8, storing 1 into
into \texttt{count} and storing \texttt{True} into \texttt{done[1]};
\item Process 0 continues execution, storing value 1 into \texttt{count}
and storing \texttt{True} into \texttt{done[0]}.
\end{itemize}

@@

\vspace{1em}
\begin{tabular}{|l|l|}
\hline
Address & compute address \\
\hline
Apply & pop $m$ and $i$ and apply $i$ to $m$, pushing a value \\
\hline
Assert & pop $b$ and check that it is \texttt{True} \\
\hline
AtomicInc & increment the atomic counter of this context \\
\hline
AtomicDec & decrement the atomic counter of this context \\
\hline
Continue & no-op (but causes a context switch) \\
\hline
Choose & choose an element from the set on top of the stack \\
\hline
Del & delete a shared variable \\
\hline
DelVar $v$ & delete process variable $v$ \\
\hline
Dict & pop $n$ and $n$ key/value pairs and push a dictionary \\
\hline
Dup & duplicate the top element of the stack \\
\hline
Frame $m$ $a$ $r$& start method $m$ with arguments $a$,
initializing variables.  Last instruction is at $r$ \\
\hline
Go & pop context and value, push value on context's stack, and add to context bag \\
\hline
Jump $p$ & set program counter to $p$ \\
\hline
JumpCond $e$ $p$ & pop expression and, if equal to $e$, set program counter to $p$ \\
\hline
Load $v$ & push shared variable $v$ (\texttt{\^} is indirect) onto the stack \\
\hline
LoadVar $v$ & load process variable $v$ and push it onto the stack \\
\hline
$n$-ary \textit{op} & apply $n$-ary operator \textit{op} to the top $n$ elements on the stack \\
\hline
Pop & pop a value of the stack and discard it \\
\hline
Push $c$ & push constant $c$ onto the stack \\
\hline
PushAddress $v$ & push the address of shared variable $v$ \\
\hline
Return & pop return address, push \texttt{result}, and restore program counter \\
\hline
Set & pop $n$ and $n$ elements and push a set of the elements \\
\hline
Spawn & pop tag, argument, and method and spawn a new context \\
\hline
Split & pop set and push minimum and remainder \\
\hline
Stop & save context into shared variable and remove from context bag \\
\hline
Store $v$ $[n]$ & pop a value from the stack and store it in shared variable $v$ (\texttt{\^} is indirect) \\
\hline
StoreVar $v$ & pop a value from the stack and store it in process variable $v$ \\
\hline
\end{tabular}
\vspace{1em}
